The function-sheaf dictionary for smooth commutative group schemes over finite fields

### 2014-08-25 crelle-dictionary.tex
Typos fixed in Applications

Suppose $X_K\in \mathcal{N}_K$ with N\'eron model $X$ and $Y_{L}\in \mathcal{N}_L$ with N\'eron model $Y$.
Suppose $m$ is a non-negative integers such that 
\[
\OK/\varpi_K^m\OK \iso \OL/\varpi_L^m\OL
\]
%(in which case $K$ and $L$ are said to be $m$-close).
Suppose also that
\begin{equation}\label{eq:schematic_transfer}
X \times_{\Spec{\OK}} \Spec{\OK/\varpi_K^m\OK} \iso Y \times_{\Spec{\OL}} \Spec{\OL/\varpi_L^m\OL}
\end{equation}
as smooth group schemes over $\OK/\varpi_K^m\OK$. 
Then
$
\Gr^{\OK}_m(X) \iso \Gr^{\OL}_{m}(Y)
$
as smooth group schemes over $\Fq$; accordingly, by Lemma~\ref{lem:pullback} the isomorphism above determines an equivalence of categories
\begin{equation}\label{eq:categorical_transfer}
\QC(\Gr^{\OK}_m(X)) \cong \QC(\Gr^{\OL}_{m}(Y))
\end{equation}
which induces an isomorphism
\begin{equation}\label{eq:transfer}
\Hom_m(X(K),\EEx)  \iso  \Hom_{m}(Y(L),\EEx).
\end{equation}
The isomorphism \eqref{eq:transfer} is an instance of {\it transfer} of (certain) quasicharacters between $X(K)$ and $Y(L)$. 
We now recognize this transfer of quasicharacters as a consequence of the equivalence of categories of quasicharacter sheaves \eqref{eq:categorical_transfer}.

The isomorphism \eqref{eq:schematic_transfer} can indeed exist between quasicharacters of algebraic tori over local fields, even when the characteristics of $K$ and $L$ differ.
Suppose $T_K$ and $M_L$ are tori over non-archimedian local fields $K$ and $L$,
splitting over $K'$ and $L'$ respectively.
Then $T_K$ and $M_L$ are said to be \emph{$n$-congruent} \cite{chai-yu:01a}*{\S 2} if there are isomorphisms
 \begin{align*}
  \alpha : \OO{K'}/\varpi_K^n \OO{K'} &\to \OO{L'}/\varpi_{L}^n \OO{L'} \\
  \beta : \Gal(K'/K) &\to \Gal(L'/L) \\
  \phi : X^*(T_K) &\to X^*(M_L)
 \end{align*}
 satisfying the conditions
 \begin{enumerate}
  \item $\alpha$ induces an isomorphism $\OK/\pi_K^n \OK \to \OO{L}/\pi_{L}^n \OO{L}$,
  \item $\alpha$ is $\Gal(K'/K)$-equivariant relative to $\beta$, and
  \item $\phi$ is $\Gal(K'/K)$-equivariant relative to $\beta$.
 \end{enumerate}
If $T_K$ and $M_L$ are $n$-congruent then $\alpha$, $\beta$ and $\phi$ determine an isomorphism 
\begin{equation}\label{transfer}
  \Hom_{n-1}(T_K(K), \EEx) \iso \Hom_{n-1}(M_L(M),\EEx).
\end{equation}
Note that if $T_K$ and $M_L$ are $n$-congruent, then they are $m$ congruent for every
$m \leq n$.
%the isomorphism \eqref{transfer} is level-preserving at integers less than $N$.
%

Now let $T$ be a N\'eron model for $T_K$ and let $M$ be a N\'eron model for $M_L$.
One of the main results of \cite{chai-yu:01a} gives an isomorphism of group schemes between the
$T \times_{\Spec{\OK}} \Spec{\OK/\varpi_n\OK}$ and $M \times_{\Spec{\OL}} \Spec{\OL/\varpi^n\OL}$, assuming that $T_K$ and $M_L$ are sufficiently congruent.
They define a quantity $h$ (the smallest integer so that $\varpi^h$ lies in the
Jacobian ideal associated to a natural embedding of $T$ into an induced torus \cite{chai-yu:01a}*{\S 8.1}) and show  that if $n > 3h$ and $T_K$ and $M_L$ are $n$-congruent then there is a canonical isomorphism of smooth group schemes
 $
\Gr_{n-3h}(T) \to \Gr_{n-3h}(M)
 $
 determined by $\alpha, \beta$ and $\phi$ \cite{chai-yu:01a}*{Thm. 8.5}.
Combining this with the paragraph above gives an instance of the geometrization and categorification of the transfer of quasicharacters.

\begin{proposition}\label{prop:transfer}
 With notation above, suppose that $T_K$ and $M_L$ are $n$-congruent and $n > 3h$.  Set $m = n-3h$.
 Then there is a canonical equivalence of categories
 \[
 \QC(\Gr^{\OK}_m(T)) \iso \QC(\Gr^{\OL}_m(M))
 \]
 determined by $\alpha, \beta$ and $\phi$ inducing an isomorphism
 \[
\Hom_m(T_K(K), \EEx) \iso  \Hom_m(M_L(L), \EEx)
 \]
compatible with \eqref{transfer}.% ($m < N$).
\end{proposition}

\acknowl{We thank Masoud Kamgarpour, Pramod Achar and Hadi Salmasian
for allowing us to hijack much of our Research in Teams meeting at BIRS into a discussion of
quasicharacter sheaves; their knowledge and help have been invaluable.
We thank Takashi Suzuki for some very helpful observations and suggestions, including those related to our use of the Hochschild-Serre spectral sequence.
We thank Joseph Bernstein for suggesting a connection between our category of quasicharacter sheaves on $G$ and his sheaves on the stacky classifying space of $G$.
We thank Alessandra Bertrapelle and Cristian Gonz\'ales-Avil\'es for disabusing us of a misapprehension concerning the Greenberg realization functor and drawing our attention to their result on Weil restriction and the Greenberg transform.
}

>>>>

Suppose $X_K\in \mathcal{N}_K$ with N\'eron model $X$ and $Y_{L}\in \mathcal{N}_L$ with N\'eron model $Y$.
Suppose $m$ is a positive integer such that 
\[
\OK/\varpi_K^{m}\OK \iso \OL/\varpi_L^{m}\OL.
\]
%(in which case $K$ and $L$ are said to be $m$-close).
Suppose also that
\begin{equation}\label{eq:schematic_transfer}
X \times_{\Spec{\OK}} \Spec{\OK/\varpi_K^{m}\OK} \iso Y \times_{\Spec{\OL}} \Spec{\OL/\varpi_L^{m}\OL}
\end{equation}
as smooth group schemes over $\OK/\varpi_K^{m}\OK$. 
Then
$
\Gr^{\OK}_{m-1}(X) \iso \Gr^{\OL}_{m-1}(Y)
$
as smooth group schemes over $\Fq$; accordingly, by Lemma~\ref{lem:pullback} the isomorphism above determines an equivalence of categories
\begin{equation}\label{eq:categorical_transfer}
\QC(\Gr^{\OK}_{m-1}(X)) \cong \QC(\Gr^{\OL}_{m-1}(Y))
\end{equation}
which induces an isomorphism
\begin{equation}\label{eq:transfer}
\Hom_{m-1}(X(K),\EEx)  \iso  \Hom_{m-1}(Y(L),\EEx).
\end{equation}
The isomorphism \eqref{eq:transfer} is an instance of {\it transfer} of (certain) quasicharacters between $X(K)$ and $Y(L)$. 
We now recognize this transfer of quasicharacters as a consequence of the equivalence of categories of quasicharacter sheaves \eqref{eq:categorical_transfer}.

The isomorphism \eqref{eq:schematic_transfer} can indeed exist between quasicharacters of algebraic tori over local fields, even when the characteristics of $K$ and $L$ differ.
Suppose $T_K$ and $M_L$ are tori over non-archimedian local fields $K$ and $L$,
splitting over $K'$ and $L'$ respectively.
Then $T_K$ and $M_L$ are said to be \emph{$n$-congruent} \cite{chai-yu:01a}*{\S 2} if there are isomorphisms
 \begin{align*}
  \alpha : \OO{K'}/\varpi_K^n \OO{K'} &\to \OO{L'}/\varpi_{L}^n \OO{L'} \\
  \beta : \Gal(K'/K) &\to \Gal(L'/L) \\
  \phi : X^*(T_K) &\to X^*(M_L)
 \end{align*}
 satisfying the conditions
 \begin{enumerate}
  \item $\alpha$ induces an isomorphism $\OK/\pi_K^n \OK \to \OO{L}/\pi_{L}^n \OO{L}$,
  \item $\alpha$ is $\Gal(K'/K)$-equivariant relative to $\beta$, and
  \item $\phi$ is $\Gal(K'/K)$-equivariant relative to $\beta$.
 \end{enumerate}
If $T_K$ and $M_L$ are $n$-congruent then $\alpha$, $\beta$ and $\phi$ determine an isomorphism 
\begin{equation}\label{transfer}
  \Hom_{n-1}(T_K(K), \EEx) \iso \Hom_{n-1}(M_L(M),\EEx).
\end{equation}
Note that if $T_K$ and $M_L$ are $n$-congruent, then they are $m$-congruent for every
$m \leq n$.
%the isomorphism \eqref{transfer} is level-preserving at integers less than $N$.
%

Now let $T$ be a N\'eron model for $T_K$ and let $M$ be a N\'eron model for $M_L$.
One of the main results of \cite{chai-yu:01a} gives an isomorphism of group schemes 
\[
T \times_{\Spec{\OK}} \Spec{\OK/\varpi_K^n\OK} \iso M \times_{\Spec{\OL}} \Spec{\OL/\varpi_L^n\OL}
\] 
assuming that $T_K$ and $M_L$ are sufficiently congruent.
They define a quantity $h$ (the smallest integer so that $\varpi^h$ lies in the
Jacobian ideal associated to a natural embedding of $T_K$ into an induced torus \cite{chai-yu:01a}*{\S 8.1}) and show  that if $n > 3h$ and $T_K$ and $M_L$ are $n$-congruent then there is a canonical isomorphism of smooth group schemes
 $
\Gr_{n-3h-1}(T) \to \Gr_{n-3h-1}(M)
 $
 determined by $\alpha, \beta$ and $\phi$ \cite{chai-yu:01a}*{Thm. 8.5}.
Combining this with the paragraph above gives the following instance of the geometrization and categorification of the transfer of quasicharacters.

\begin{proposition}\label{prop:transfer}
 With notation above, suppose that $T_K$ and $M_L$ are $n$-congruent and $n > 3h$.  Set $m = n-3h$.
 Then there is a canonical equivalence of categories
 \[
 \QC(\Gr^{\OK}_{m-1}(T)) \iso \QC(\Gr^{\OL}_{m-1}(M))
 \]
 determined by $\alpha, \beta$ and $\phi$ inducing an isomorphism
 \[
\Hom_{m-1}(T_K(K), \EEx) \iso  \Hom_{m-1}(M_L(L), \EEx)
 \]
compatible with \eqref{transfer}.% ($m < N$).
\end{proposition}

\acknowl{We thank Masoud Kamgarpour, Pramod Achar and Hadi Salmasian
for allowing us to hijack much of our Research in Teams meeting at BIRS into a discussion of
quasicharacter sheaves; their knowledge and help have been invaluable.
We thank Takashi Suzuki for some very helpful observations and suggestions, especially related to our use of the Hochschild-Serre spectral sequence.
We thank Joseph Bernstein for suggesting a connection between our category of quasicharacter sheaves for $p$-adic tori and quasicharacters of its pure rational forms.
We thank Alessandra Bertrapelle and Cristian Gonz\'ales-Avil\'es for disabusing us of a misapprehension concerning the Greenberg realization functor and drawing our attention to their result on Weil restriction and the Greenberg transform.
}
